---
title: "Homework5Part1" 
output: pdf_document
author: "Rafael Espinosa Castaneda"
date: "2023-02-18"
---

```{r setup, include=TRUE,message=FALSE,warning=FALSE}
# --------------------------Setting the parameters to be used------------------------
###################################################
## COMPUTING DEMO:  Kriging
##
## DSC 383: Advanced Predictive Models for Complex Data
## By:  Kate Calder, UT Austin
## Last updated: September 26, 2021
###################################################

## load libraries
library(ggplot2)
library(mapproj)
library(geoR)
library(leaflet)
library(gridExtra)
library(measurements)
library(dplyr)


# load scallops data
rain <- read.table("/Users/rafa/Documents/Master Austin/MAESTRIA_AUSTIN/Advanced Predictive Models/Hw5/rain.txt",header=T)

rain$altitude<-conv_unit(rain$altitude, "ft", "mi") 




```

## Part a

```{r parta, include=TRUE,message=FALSE,warning=FALSE}

# create geodata (geoR) object
rain_geo_altitude <- as.geodata(
  cbind(rain$altitude, 
        rain$x, 
        rain$y), 
  data.col = 1, 
  coords.col = 2:3)



# create geodata (geoR) object
rain_geo_rainfall <- as.geodata(
  cbind(rain$rainfall, 
        rain$x, 
        rain$y), 
  data.col = 1, 
  coords.col = 2:3)


par(mfrow=c(1,2))

points.geodata(rain_geo_rainfall, col = gray(seq(1, 0.1, l = 100)),xlab = "x", ylab = "y")
title("Spatial variation in rainfall ")

points.geodata(rain_geo_altitude, col = gray(seq(1, 0.1, l = 100)),xlab = "x", ylab = "y")
title("Spatial variation in altitude ")

```
$$\textbf{ANSWER}$$

We can notice certain spatial dependence for the altitude. We can notice from the plot that in general the altitude is greater for coordinates $x>=100$ and $y<150$. For the rainfall variable we can notice a certain dependece such that greater values are around the center of the 2D coordinates.



## Part b

```{r partb, include=TRUE,message=FALSE,warning=FALSE}
lR_RA = lm(sqrt(rainfall)~altitude, data = rain) 

summary(lR_RA)

print("The estimated error variance")
print(4.275^2)

print("The proportion of variance is ")

print(0.03145)

```

$$\textbf{ANSWER}$$

The equation is  $\sqrt{rainfall}=14.1273-8.3681*Altitude$. The estimated error variance is $18.27563$ and the proportion of variation in the square root of rainfall is explained by altitude is
$0.03145$


## Part c


```{r partc, include=TRUE,message=FALSE,warning=FALSE}

N<-length(rain$rainfall)
#tot<-N*(N-1)

tot<-N*(N-1)/2
print(N)
print(tot)
distances<-rep(1,tot)
count<-0
for (i in 1:N)
     {
       for(j in (i+1):N)
       {
         if (i!=j)
         {
           count<-count+1
         distances[count]<- ( (rain$x[i]-rain$x[j])^2 +(rain$y[i]-rain$y[j])^2 )^(1/2)
         
         if (is.na(distances[count]))
         {
           print(paste("i",i,"j",j))
           print("The x_i")
           print(rain$x[i])
           print("The x_j")
           print(rain$x[j])
           print("The y_i")
           print(rain$y[i])
           print("The y_j")
           print(rain$y[j])
         }
         
         
         
         }
         
        }
       
     }
hist(distances, breaks=seq(from=0, to=300, by=20),freq=FALSE,xlab="Distances",ylab="Relative Frequency")

```

## Part d

```{r partd, include=TRUE,message=FALSE,warning=FALSE}

residuals<-resid(lR_RA)

bins_ranges<-c(20,40,60,80,100,120,140,160,180,200,220,240,260,280,300)






df1 <- data.frame(x_1=c(rain$x[1]),
                  x_2=c(rain$x[2]),distance=c( ( (rain$x[1]-rain$x[2])^2 +(rain$y[1]-rain$y[2])^2 )^(1/2) ) , res1=c (residuals[1]), res2=c (residuals[2]))


for (i in 1:N)
  {
  for(j in (i+1):N)
    {
    if (i!=j && (i!=1 && j!=2)  )
      {
      count<-count+1
      distances[count]<- ( (rain$x[i]-rain$x[j])^2 +(rain$y[i]-rain$y[j])^2 )^(1/2)
      
      
          df2 <- data.frame(x_1=c(rain$x[i]),
                            x_2=c(rain$x[j]),distance=c( ( (rain$x[i]-rain$x[j])^2 +(rain$y[i]-rain$y[j])^2 )^(1/2) ) , res1=c (residuals[i]), res2=c (residuals[j]))
          
          df1<-bind_rows(df1, df2)
          
      }
    
  }

  
  }


correlation<- rep(0,length(bins_ranges))
counters_dist<- rep(0,length(bins_ranges))

count<-1

for (bin in bins_ranges)
{
  df_to_use<-filter(df1, distance<bin & distance>=bin-20 & x_1 !=x_2)
  counters_dist[count]<-length(df_to_use$x_1)
  correlation[count]<-cor(df_to_use$res1,df_to_use$res2)
  count<-count+1
}


df_to_use

centers<-c(10,30,50,70,90,110,130,150,170,190,210,230,250,270,290)



df2<-data.frame(centerBins=centers,corr=correlation,num_dist=counters_dist)


counters_dist
as_geo_corr <- as.geodata(
  cbind(counters_dist, 
        centers, 
        correlation), 
  data.col = 1, 
  coords.col = 2:3)

#points.geodata(as_geo_corr,xlab = "center bins", ylab = "correlation")


ggplot() + geom_point(data = df2, aes(x =centerBins , y = corr),cex=10*(counters_dist+min(counters_dist))/max(counters_dist)+2.5)+labs(x = "Distance",y = "Correlation")



```

## Part e

In Switzerland, a small country it shouldn't be very likely to have pair of locations extremely far one from the other. That's why it is much more likely a distance between 80 and 100 than distances greater than 280.  
